Synthesising Recursive Functions for\\ First-Order Model Counting

### 2023-03-30 doc/talks/3_long/talk.tex
Finish the introduction

\author{Paulius Dilkas \\ \vspace{1cm} \tiny{Joint work with Vaishak Belle (Univeristy of Edinburgh, UK)}}
\title{Synthesising Recursive Functions for\\ First-Order Model Counting}
\institute{National University of Singapore, Singapore}
\date{\today}
% TODO: update the date
% TODO: unify the use of colours
% TODO: do I want to mention forclift and/or crane?
% TODO: maybe replace red!50 with one of the main colours
% TODO: \color -> \textcolor everywhere
% TODO: how do we capitalise blocks?
% TODO: make sure I'm using the 'forest' package whenever possible
% TODO: use Latex and not latex everywhere

@@

% TODO: add a slide about other approaches to WFOMC, making it crystal clear
% what my contributions are. This will finish the introduction

@@

@@

\begin{frame}{First-Order Knowledge Compilation: An Approach to FOMC}
  \begin{minipage}[t][0.81\textheight][t]{\textwidth}
    \centering
    \begin{forest}
      for tree={s sep=10mm, sn edges}
      [\alt<3->{$\bigwedge_{c \in \Delta}$}{$\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$},
      inner on=<3->,
      label={[visible on=<{13-14,19-}>]right:{\alert{\alt<15->{$3^{|\Delta|}$}{$\{\,\pP(c), \qQ(c)\,\}$}}}}
      [\alt<5->{$\lor$}{$\pP(c) \lor \qQ(c)$},
      name=lor,
      for tree={visible on=<3->},
      inner on=<5->,
      label={[visible on=<{12-14,18-}>]right:{\alert{\only<18->{$2 + 1 = 3$}\only<12-13>{\colorbox{white}{$\{\,\pP(c), \qQ(c)\,\}$}}\only<14>{\colorbox{redish}{$\{\,\pP(c), \qQ(c)\,\}$}}}}}
      [\alt<7->{$\land$}{$\begin{aligned}\qQ(c)\\ \pP(c) \lor \qQ(c)\end{aligned}$},
      name=and,
      for tree={visible on=<5->},
      inner on=<7->,
      label={[visible on=<{11-14,17-}>]left:{\alert{\only<17->{$2 \times 1 = 2$}\only<11-13>{\colorbox{white}{$\{\,\qQ(c)\,\}$}}\only<14>{\colorbox{redish}{$\{\,\qQ(c)\,\}$}}}}}
      [\alt<15->{$\pP(c) \lor \neg \pP(c)$}{$\qQ(c)$},
      name=smoothing,
      for tree={visible on=<7->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{2}}}]
      [\alt<15->{$\land$}{$\top$},
      for tree={visible on=<7->},
      leaf on=<6-14>,
      inner on=<15->,
      label={[label distance=0cm,visible on=<17->]87:{\alert{$1 \times 1 = 1$}}},
      label={[visible on=<10>]below:{\alert{tautology}}},
      [$\qQ(c)$,
      for tree={visible on=<15->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      [$\top$,
      for tree={visible on=<15->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      ]
      ]
      [\alt<9->{$\land$}{$\begin{aligned}\neg\qQ(c)\\ \pP(c) \lor \qQ(c)\end{aligned}$},
      name=negformula,
      for tree={visible on=<5->},
      inner on=<9->,
      label={[visible on=<{11-14,17-}>]right:{\alert{\alt<15->{$1 \times 1 = 1$}{$\{\,\pP(c), \qQ(c)\,\}$}}}}
      [$\neg \qQ(c)$,
      name=negq,
      for tree={visible on=<9->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      [$\pP(c)$,
      name=pnode,
      for tree={visible on=<9->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      ]
      ]
      ]
      \begin{scope}[on background layer]
        \uncover<6>{\node[fill=redish, ellipse, fit={(and)}, inner sep=0.5pt] {};}
        \uncover<8>{\node[fill=redish, ellipse, fit={(negformula)}, inner sep=0.5pt] {};}
        \uncover<15>{\node[fill=redish, ellipse, fit={(smoothing) (and)}, inner sep=0.5pt] {};}
      \end{scope}
      \node<10>[below = 3cm of lor] (n1) {\alert{unit clauses}};
      \path[-Latex,draw]<10> (n1) edge[bend right] (pnode);
      \path[-Latex,draw]<10> (n1) edge (negq);
      \path[-Latex,draw]<10> (n1) edge[bend left] (smoothing);
    \end{forest}
  \end{minipage}
  \begin{minipage}{\textwidth}
    \begin{overlayarea}{\textwidth}{4cm}
      \only<2-3>{\centerline{Independent partial grounding (introduces a constant \alert{$c \in \Delta$})}}%
      \only<4-5>{\centerline{Shannon decomposition (a.k.a. Boole's expansion theorem) on \alert{$\qQ(c)$}}}%
      \only<6-7>{\centerline{Positive unit propagation of \alert{$\qQ(c)$}}}%
      \only<8-9>{\centerline{Negative unit propagation of \alert{$\neg\qQ(c)$}}}%
      \only<10>{\centerline{Compilation is complete \textcolor{forestgreen}{\cmark}}}%
      \only<11-13>{\centerline{Smoothing: propagating atoms upwards}}%
      \only<14-15>{\centerline{Smoothing: adding new atoms}}%
      \only<16->{\centerline{Propagating the model count}}%
    \end{overlayarea}
  \end{minipage}
\end{frame}

@@

\begin{frame}{Compilation Rules}
  \begin{definition}
    A \alert{(compilation) rule} is a function that takes a \alert{formula} and returns a set of \alert{$(G, L)$} pairs, where
    \begin{itemize}
    \item \alert{$G$} is an FCG,
    \item and \alert{$L$} is a list of formulas.
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}{Example Rule: Independence}
  Input formula:
  \begin{gather}
    ({\color{color1} \forall x, y \in \Omega\text{. }x=y}) \land \label{eq:1} \tag{{\color{color1} 1}} \\
    ({\color{color2} \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\texttt{P}(x, y) \land \texttt{P}(x, z) \Rightarrow y=z}) \land \label{eq:2} \tag{{\color{color2} 2}} \\
    ({\color{color2} \forall w, x \in \Gamma\text{. }\forall y \in \Delta\text{. }\texttt{P}(w, y) \land \texttt{P}(x, y) \Rightarrow w=x}) \label{eq:3} \tag{{\color{color2} 3}}
  \end{gather}
  \pause
  Only one \alert{(G, L)} pair:
  \[
  G =
  \begin{tikzpicture}[edge from parent/.style={draw,-latex},baseline=(current bounding box.center)]
    \node[draw,circle] {$\land$}
    child {node {\color{color1} $\star$}}
    child {node {\color{color2} $\star$}}
    ;
  \end{tikzpicture},
  \qquad
  L = \langle {\color{color1} \eqref{eq:1}}, {\color{color2} \eqref{eq:2} \land \eqref{eq:3}} \rangle
  \]
\end{frame}
% stars mark edges to nowhere and they are associated with the formulas in L via a certain order

\begin{frame}
  \frametitle<1>{New Rule 1: Generalised Domain Recursion}
  \frametitle<2>{New Rule 2: Constraint Removal}
  \frametitle<3>{New Rule 3: Identifying Possibilities for Recursion}
  \begin{columns}
    \begin{column}{0.15\textwidth}
      \centering
      \begin{tikzpicture}[every node/.style={draw,ellipse},edge from parent/.style={draw,-latex},sibling distance=10mm,level distance=10mm]
        \node[fill=red!50, fill on=<1>] (dr) {}
        child {node {}
          child {node[fill=red!50, fill on=<2>] {}
            child {node {}
              child {node {}}
              child {node[fill=red!50, fill on=<3>] (ref) {}}
            }}};
        \draw[-latex, bend right,alt=<3>{red}{black}] (ref) to (dr);
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.85\textwidth}
      \begin{overprint}
        \onslide<1>
          \begin{example}
            Input formula:
            \[
              \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
            \]
            Output formula (with a new constant \alert{$c \in \Gamma$}):
            \begin{gather*}
              \forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(\alert{c}, y) \lor \neg \texttt{P}(\alert{c}, z) \\
              \begin{multlined}
                \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\alert{x \ne c} \land y \ne z \Rightarrow \\
                \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
              \end{multlined}
            \end{gather*}
            %% Output FCG:
            %% \begin{tikzpicture}[edge from parent/.style={draw,-latex}]
            %%   \node[draw,circle] {$\textsc{DR}$}
            %%   child {node {$\star$}}
            %%   ;
            %% \end{tikzpicture}
          \end{example}
        \onslide<2>
          \begin{example}
            Input formula (with a constant \alert{$c \in \Gamma$}):
            \begin{gather*}
              \begin{multlined}
                \forall x \in \alert{\Gamma}\text{. }\forall y, z \in \Delta\text{. }\alert{x \ne c} \land y \ne z \Rightarrow \\
                \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
              \end{multlined} \\
              \begin{multlined}
                \forall w,x \in \alert{\Gamma}\text{. }\forall y \in \Delta\text{. }\alert{w \ne c} \land \alert{x \ne c} \land w \ne x \Rightarrow \\
                \neg \texttt{P}(w, y) \lor \neg \texttt{P}(x, y)
              \end{multlined}
            \end{gather*}
            Output formula (with a new domain \alert{$\Gamma' \coloneqq \Gamma \setminus \{\, c \,\}$}):
            \begin{gather*}
              \forall x \in \alert{\Gamma'}\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z) \\
              \forall w,x \in \alert{\Gamma'}\text{. }\forall y \in \Delta\text{. }w \ne x \Rightarrow \neg \texttt{P}(w, y) \lor \neg \texttt{P}(x, y)
            \end{gather*}
          \end{example}
        \onslide<3>
          \begin{block}{Goal}
            Check if the input formula is isomorphic (up to domains) to a
            previously encountered formula.
          \end{block}
          \begin{block}{Rough Outline}
            \begin{enumerate}
              \item Consider pairs of `similar' clauses.
              \item Consider bijections between their sets of variables.
              \item Extend each such bijection to a map between sets of domains.
              \item If the bijection makes the clauses equal, and the domain map
                    is compatible with previous domain maps, move on to another
                    pair of clauses.
            \end{enumerate}
          \end{block}
      \end{overprint}
    \end{column}
  \end{columns}
\end{frame}

@@

  Our new rules enable \textsc{ForcLift} to efficiently count
  \alert{$\Gamma \to \Delta$} functions such as:
  \begin{itemize}
    \item injections in \alert{$\Theta(mn)$} time
          \begin{itemize}
            \item best: \alert{$\Theta(m)$}
          \end{itemize}
    \item partial injections in \alert{$\Theta(mn)$} time
          \begin{itemize}
            \item best: \alert{$\Theta({\min\{\, m, n \,\}}^2)$}
          \end{itemize}
          \item bijections in \alert{$\Theta(m)$} time
          \begin{itemize}
            \item \alert{optimal!}
          \end{itemize}
  \end{itemize}
\end{frame}
% TODO: mention Crane, don't forget to introduce forclift
% TODO: mention that everything used to be unliftable (by any algorithm with a publicly available implementation)
% TODO: what font do I want to use for forclift/crane? It's mentioned on a slide title as well

% TODO: maybe replace the zoom picture with some galaxy picture?
{ % all template changes are local to this group.
    \setbeamertemplate{navigation symbols}{}
    \begin{frame}<article:0>[plain]
        \begin{tikzpicture}[remember picture,overlay]
            \node[at=(current page.center)] {
                \includegraphics[keepaspectratio, height=\paperheight]{zoom.jpg}
            };
        \end{tikzpicture}
     \end{frame}
}

>>>>

\author{Paulius Dilkas \\ \vspace{1cm} \tiny{Joint work with Vaishak Belle (Univeristy of Edinburgh, UK)}}
\title{Synthesising Recursive Functions for\\ First-Order Model Counting}
\institute{National University of Singapore, Singapore}
\date{\today}
% TODO: update the date
% TODO: unify the use of colours
% TODO: maybe replace red!50 with one of the main colours
% TODO: \color -> \textcolor everywhere
% TODO: how do we capitalise titles of \block?
% TODO: make sure I'm using the 'forest' package whenever possible
% TODO: use Latex and not latex everywhere

@@

@@

\begin{frame}{Approaches to FOMC}
  \begin{itemize}
    \item \alert{ForcLift}~\textcolor{gray}{\parencite{DBLP:conf/ijcai/BroeckTMDR11}}
    \begin{itemize}
      \item knowledge compilation to \alert{FO d-DNNF}
    \end{itemize}
    \item \alert{L2C}~\textcolor{gray}{\parencite{DBLP:conf/kr/KazemiP16}}
    \begin{itemize}
      \item knowledge compilation to \alert{C++} code
    \end{itemize}
    \item \alert{Alchemy}~\textcolor{gray}{\parencite{DBLP:journals/cacm/GogateD16}}
    \begin{itemize}
      \item \alert{DPLL}-style search
    \end{itemize}
    \item \alert{FastWFOMC}~\textcolor{gray}{\parencite{DBLP:conf/uai/BremenK21}}
    \begin{itemize}
      \item knowledge compilation to \alert{sd-DNNF}
    \end{itemize}
  \end{itemize}
  \pause
  \vfill
  \begin{block}{Our Contribution}
    \centering
    \begin{tikzpicture}[align=center,node distance=0.1cm]
      \node[label=below:{ForcLift}] (forclift) {\includegraphics[width=0.2\paperwidth]{forklift.png}};
      \node[right=of forclift] (plus) {\Huge$+$};
      \node[right=of plus,label=below:{Recursion}] (recursion) {\includegraphics[width=0.2\paperwidth]{matryoshka.jpg}};
      \node[right=of recursion] (equal) {\Huge$=$};
      \node[right=of equal,label=below:{Crane}] (crane) {\includegraphics[width=0.2\paperwidth]{crane.jpg}};
    \end{tikzpicture}
  \end{block}
\end{frame}

@@

\begin{frame}{ForcLift and First-Order Knowledge Compilation}
  \begin{minipage}[t][0.81\textheight][t]{\textwidth}
    \centering
    \begin{forest}
      for tree={s sep=10mm, sn edges}
      [\alt<3->{$\bigwedge_{c \in \Delta}$}{$\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$},
      inner on=<3->,
      label={[visible on=<{13-14,19-}>]right:{\alert{\alt<15->{$3^{|\Delta|}$}{$\{\,\pP(c), \qQ(c)\,\}$}}}}
      [\alt<5->{$\lor$}{$\pP(c) \lor \qQ(c)$},
      name=lor,
      for tree={visible on=<3->},
      inner on=<5->,
      label={[visible on=<{12-14,18-}>]right:{\alert{\only<18->{$2 + 1 = 3$}\only<12-13>{\colorbox{white}{$\{\,\pP(c), \qQ(c)\,\}$}}\only<14>{\colorbox{redish}{$\{\,\pP(c), \qQ(c)\,\}$}}}}}
      [\alt<7->{$\land$}{$\begin{aligned}\qQ(c)\\ \pP(c) \lor \qQ(c)\end{aligned}$},
      name=and,
      for tree={visible on=<5->},
      inner on=<7->,
      label={[visible on=<{11-14,17-}>]left:{\alert{\only<17->{$2 \times 1 = 2$}\only<11-13>{\colorbox{white}{$\{\,\qQ(c)\,\}$}}\only<14>{\colorbox{redish}{$\{\,\qQ(c)\,\}$}}}}}
      [\alt<15->{$\pP(c) \lor \neg \pP(c)$}{$\qQ(c)$},
      name=smoothing,
      for tree={visible on=<7->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{2}}}]
      [\alt<15->{$\land$}{$\top$},
      for tree={visible on=<7->},
      leaf on=<6-14>,
      inner on=<15->,
      label={[label distance=0cm,visible on=<17->]87:{\alert{$1 \times 1 = 1$}}},
      label={[visible on=<10>]below:{\alert{tautology}}},
      [$\qQ(c)$,
      for tree={visible on=<15->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      [$\top$,
      for tree={visible on=<15->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      ]
      ]
      [\alt<9->{$\land$}{$\begin{aligned}\neg\qQ(c)\\ \pP(c) \lor \qQ(c)\end{aligned}$},
      name=negformula,
      for tree={visible on=<5->},
      inner on=<9->,
      label={[visible on=<{11-14,17-}>]right:{\alert{\alt<15->{$1 \times 1 = 1$}{$\{\,\pP(c), \qQ(c)\,\}$}}}}
      [$\neg \qQ(c)$,
      name=negq,
      for tree={visible on=<9->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      [$\pP(c)$,
      name=pnode,
      for tree={visible on=<9->},
      rectangle,
      draw,
      fill=gray!25,
      label={[visible on=<16->]below:{\alert{1}}}]
      ]
      ]
      ]
      \begin{scope}[on background layer]
        \uncover<6>{\node[fill=redish, ellipse, fit={(and)}, inner sep=0.5pt] {};}
        \uncover<8>{\node[fill=redish, ellipse, fit={(negformula)}, inner sep=0.5pt] {};}
        \uncover<15>{\node[fill=redish, ellipse, fit={(smoothing) (and)}, inner sep=0.5pt] {};}
      \end{scope}
      \node<10>[below = 3cm of lor] (n1) {\alert{unit clauses}};
      \path[-Latex,draw]<10> (n1) edge[bend right] (pnode);
      \path[-Latex,draw]<10> (n1) edge (negq);
      \path[-Latex,draw]<10> (n1) edge[bend left] (smoothing);
    \end{forest}
  \end{minipage}
  \begin{minipage}{\textwidth}
    \begin{overlayarea}{\textwidth}{4cm}
      \only<2-3>{\centerline{Independent partial grounding (introduces a constant \alert{$c \in \Delta$})}}%
      \only<4-5>{\centerline{Shannon decomposition (a.k.a. Boole's expansion theorem) on \alert{$\qQ(c)$}}}%
      \only<6-7>{\centerline{Positive unit propagation of \alert{$\qQ(c)$}}}%
      \only<8-9>{\centerline{Negative unit propagation of \alert{$\neg\qQ(c)$}}}%
      \only<10>{\centerline{Compilation is complete \textcolor{forestgreen}{\cmark}}}%
      \only<11-13>{\centerline{Smoothing: propagating atoms upwards}}%
      \only<14-15>{\centerline{Smoothing: adding new atoms}}%
      \only<16->{\centerline{Propagating the model count}}%
    \end{overlayarea}
  \end{minipage}
\end{frame}

@@

\begin{frame}{How FCGs Are Built}
  \begin{definition}
    A \alert{(compilation) rule} is a function that takes a \alert{formula} and returns a set of \alert{$(G, L)$} pairs, where
    \begin{itemize}
    \item \alert{$G$} is an FCG,
    \item and \alert{$L$} is a list of formulas.
    \end{itemize}
  \end{definition}
  \vfill The formulas in \alert{$L$} are then \alert{compiled}, and the
  resulting FCGs are \alert{inserted} into \alert{$G$} according to a \alert{set
    order}.
\end{frame}

\begin{frame}{Example Compilation Rule: Independence}
  Input formula:
  \begin{gather}
    ({\color{color1} \forall x, y \in \Omega\text{. }x=y}) \land \label{eq:1} \tag{{\color{color1} 1}} \\
    ({\color{color2} \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\texttt{P}(x, y) \land \texttt{P}(x, z) \Rightarrow y=z}) \land \label{eq:2} \tag{{\color{color2} 2}} \\
    ({\color{color2} \forall w, x \in \Gamma\text{. }\forall y \in \Delta\text{. }\texttt{P}(w, y) \land \texttt{P}(x, y) \Rightarrow w=x}) \label{eq:3} \tag{{\color{color2} 3}}
  \end{gather}
  \pause
  Only one \alert{(G, L)} pair:
  \[
  G =
  \begin{tikzpicture}[edge from parent/.style={draw,-latex},baseline=(current bounding box.center)]
    \node[draw,circle] {$\land$}
    child {node {\color{color1} $\star$}}
    child {node {\color{color2} $\star$}}
    ;
  \end{tikzpicture},
  \qquad
  L = \langle {\color{color1} \eqref{eq:1}}, {\color{color2} \eqref{eq:2} \land \eqref{eq:3}} \rangle
  \]
\end{frame}
% stars mark edges to nowhere and they are associated with the formulas in L via a certain order

\begin{frame}
  \frametitle<1>{New Rule 1: Generalised Domain Recursion}
  \frametitle<2>{New Rule 2: Constraint Removal}
  \frametitle<3>{New Rule 3: Identifying Possibilities for Recursion}
  \begin{columns}
    \begin{column}{0.15\textwidth}
      \centering
      \begin{tikzpicture}[every node/.style={draw,ellipse},edge from parent/.style={draw,-latex},sibling distance=10mm,level distance=10mm]
        \node[fill=red!50, fill on=<1>] (dr) {}
        child {node {}
          child {node[fill=red!50, fill on=<2>] {}
            child {node {}
              child {node {}}
              child {node[fill=red!50, fill on=<3>] (ref) {}}
            }}};
        \draw[-latex, bend right,alt=<3>{red}{black}] (ref) to (dr);
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.85\textwidth}
      \begin{overprint}
        \onslide<1>
          \begin{example}
            Input formula:
            \[
              \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
            \]
            Output formula (with a new constant \alert{$c \in \Gamma$}):
            \begin{gather*}
              \forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(\alert{c}, y) \lor \neg \texttt{P}(\alert{c}, z) \\
              \begin{multlined}
                \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }\alert{x \ne c} \land y \ne z \Rightarrow \\
                \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
              \end{multlined}
            \end{gather*}
            %% Output FCG:
            %% \begin{tikzpicture}[edge from parent/.style={draw,-latex}]
            %%   \node[draw,circle] {$\textsc{DR}$}
            %%   child {node {$\star$}}
            %%   ;
            %% \end{tikzpicture}
          \end{example}
        \onslide<2>
          \begin{example}
            Input formula (with a constant \alert{$c \in \Gamma$}):
            \begin{gather*}
              \begin{multlined}
                \forall x \in \alert{\Gamma}\text{. }\forall y, z \in \Delta\text{. }\alert{x \ne c} \land y \ne z \Rightarrow \\
                \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z)
              \end{multlined} \\
              \begin{multlined}
                \forall w,x \in \alert{\Gamma}\text{. }\forall y \in \Delta\text{. }\alert{w \ne c} \land \alert{x \ne c} \land w \ne x \Rightarrow \\
                \neg \texttt{P}(w, y) \lor \neg \texttt{P}(x, y)
              \end{multlined}
            \end{gather*}
            Output formula (with a new domain \alert{$\Gamma' \coloneqq \Gamma \setminus \{\, c \,\}$}):
            \begin{gather*}
              \forall x \in \alert{\Gamma'}\text{. }\forall y, z \in \Delta\text{. }y \ne z \Rightarrow \neg \texttt{P}(x, y) \lor \neg \texttt{P}(x, z) \\
              \forall w,x \in \alert{\Gamma'}\text{. }\forall y \in \Delta\text{. }w \ne x \Rightarrow \neg \texttt{P}(w, y) \lor \neg \texttt{P}(x, y)
            \end{gather*}
          \end{example}
        \onslide<3>
          \begin{block}{Goal}
            Check if the input formula is equivalent (up to domains) to a
            previously encountered formula.
          \end{block}
          \begin{block}{Outline}
            \begin{enumerate}
              \item Consider pairs of `similar' clauses.
              \item Consider bijections between their sets of variables.
              \item Extend each such bijection to a map between sets of domains.
              \item If the bijection makes the clauses equal, and the domain map
                    is compatible with previous domain maps, move on to another
                    pair of clauses.
            \end{enumerate}
          \end{block}
      \end{overprint}
    \end{column}
  \end{columns}
\end{frame}

@@

  Our new rules enable Crane to efficiently count
  \alert{$\Gamma \to \Delta$} functions such as:
  \begin{itemize}
    \item injections in \alert{$\Theta(mn)$} time
          \begin{itemize}
            \item by hand: \alert{$\Theta(m)$}
          \end{itemize}
    \item partial injections in \alert{$\Theta(mn)$} time
          \begin{itemize}
            \item by hand: \alert{$\Theta({\min\{\, m, n \,\}}^2)$}
          \end{itemize}
          \item bijections in \alert{$\Theta(m)$} time
          \begin{itemize}
            \item \alert{optimal!}
          \end{itemize}
  \end{itemize}
\end{frame}

% TODO: add a slide with the runtime plot

{ % all template changes are local to this group.
    \setbeamertemplate{navigation symbols}{}
    \begin{frame}<article:0>[plain]
        \begin{tikzpicture}[remember picture,overlay]
            \node[at=(current page.center)] {
                \includegraphics[keepaspectratio, width=\paperwidth]{hubble.jpg}
            };
            \node[shift=({-0.7, 0.1})] at (current page.south east) {\tiny \color{lightgray}{Source: NASA}};
        \end{tikzpicture}
     \end{frame}
}
